Creating the ``r params$package_name`` R package

### 2025-05-31 create-rhello.Rmd
inprod test

@@

- The `predictor_hybrid` class is an S3 object designed to represent this structure.

- We implement Hilbert space operations for this class, and our algorithm is built on these foundational operations.

The class definition is a simple list that contains a matrix and a collection of \code{fd} objects from the \code{fda} package, along with the necessary metadata.

```{r}
#' Create a predictor_hybrid object (S3 version, automatic basis size and Gram matrix)
#'
#' Constructs a hybrid predictor object that stores both scalar and functional predictors,
#' and automatically computes Gram matrices (inner products of basis functions).
#'
#' @param Z A numeric matrix of dimension \code{n_sample × n_scalar} representing scalar predictors.
#' @param functional_list A list of functional predictors (e.g., \code{fd} objects from the \code{fda} package).
#'
#' @return An object of class \code{predictor_hybrid}, containing scalar and functional data with Gram matrices.
#' @export
predictor_hybrid <- function(Z, functional_list) {
  stopifnot(is.matrix(Z), is.numeric(Z))
  stopifnot(is.list(functional_list))

@@

  structure(
    list(
      Z = Z,
      functional_list = functional_list,
      gram_list = gram_list,
      n_basis_list = n_basis_list,
      n_sample = n_sample,
      n_functional = n_functional,
      n_scalar = n_scalar
    ),
    class = "predictor_hybrid"
  )
}

Let's write some tests to make sure the function behaves as desired:
```{r}
testthat::test_that("predictor_hybrid constructor works as expected (S3, updated)", {
  library(fda)

@@

  # Check each functional component's coefficients
  expected_coef <- matrix(8, nrow = 5, ncol = 3)
  for (fd_obj in scaled$functional_list) {
    testthat::expect_equal(
      unname(coef(fd_obj)),
      unname(expected_coef)
    )
  }
})
```
### inprod.predictor_hybrid
The inner product of $f_1=(f_1^{(1)}, \ldots,f_1^{(K)}) $ and $f_2=(f_2^{(1)}, \ldots,f_2^{(K)})$ in $\mathcal{F}$ is defined as $\langle f_1, f_2\rangle_\mathcal{F} =  \sum_{k=1}^K \langle f_1^{(k)}, f_2^{(k)}\rangle_{L^2} = \sum_{k=1}^K \int_{\mathcal{T}_k} f_1^{(k)}(t_k) f_2^{(k)}(t_k) dt_k$ with norm $\Vert f_1 \Vert_\mathcal{F} = \langle f_1,f_1 \rangle_\mathcal{F}^{1/2} = \{ \sum_{k=1}^K \int_{\mathcal{T}_k} f_1^{(k)}(t_k)^2 dt_k\}^{1/2}$.

Computes the inner product between two `predictor_hybrid` objects.Both scalar and functional components are included.

  We define the inner product between any two hybrid objects, $\mathbf{h}_1 = (f_1, \mathbf{v}_1)$ and $\mathbf{h}_2 = (f_2, \mathbf{v}_2)$, as
$$
    \langle \mathbf{h}_1, \mathbf{h}_2\rangle_{\mathcal{H}} = \langle f_1, f_2\rangle_\mathcal{F} + w \langle  \mathbf{v_1}, \mathbf{v_2} \rangle = \sum \limits_{k=1}^K \int_{\mathcal{T}_k} f_1^{(k)}(t_k) f_2^{(k)}(t_k) dt_k + \omega \sum \limits_{r=1}^p v_{1r}v_{2r}, 
$$
with norm $\Vert \cdot \Vert_\mathcal{H} = \langle \cdot,\cdot \rangle_\mathcal{H}^{1/2}$. $\omega$ is a positive weight that needs to be pre-specified or estimated. It is mainly used to take into account heterogeneity between functional and scalar parts in terms of measurement scale and/or amount of variation (see Section \ref{subsec: Data Preprocessing}). Without loss of generality and for the clarity of illustration, all the following theoretical results will be derived for $\omega=1$. The results remain valid for any positive weights.
- Usage

@@

  - `xi_1`: A `predictor_hybrid` object.
  - `xi_2`: Another `predictor_hybrid` object.

- Value:  
A numeric matrix of inner products (or a scalar if both inputs are single observations).

- Details:  
  - Functional components are summed using `inprod()` from the `fda` package.  
  - Scalar components are handled via matrix multiplication.  
  - Returns a scalar if both `xi_1` and `xi_2` are single-sample predictors.
```{r}
#' Inner product between two predictor_hybrid objects (with broadcasting)
#'
#' Computes the inner product between two `predictor_hybrid` objects,
#' including both functional and scalar components. Supports broadcasting
#' when one of the inputs has a single observation. If only one input is given,
#' the inner product is computed with itself.
#'
#' @param xi_1 A `predictor_hybrid` object.
#' @param xi_2 Another `predictor_hybrid` object. If missing, defaults to `xi_1`.
#'
#' @return A numeric matrix of inner products (or a scalar if both inputs have one row).
#' @export
inprod.predictor_hybrid <- function(xi_1, xi_2 = NULL) {
  if (is.null(xi_2)) xi_2 <- xi_1

  # Optional: Check for basis compatibility
  # if (!is_same_basis(xi_1, xi_2)) stop("Functional predictors must share the same basis.")

@@

  # Compute sum of pairwise functional inner products
  inprod_functional <- Reduce(`+`,
    Map(function(f1, f2) fda::inprod(f1, f2),
        xi_1$functional_list, xi_2$functional_list)
  )

  # Broadcast-aware scalar component
  Z1 <- if (n1 == 1) matrix(xi_1$Z, nrow = 1) else xi_1$Z
  Z2 <- if (n2 == 1) matrix(xi_2$Z, nrow = 1) else xi_2$Z
  inprod_scalar <- Z1 %*% t(Z2)

  result <- inprod_functional + inprod_scalar

  if (n1 == 1 && n2 == 1) as.numeric(result) else result
}

```{r}
testthat::test_that("inprod.predictor_hybrid works with broadcasting and self-inner product", {
  suppressPackageStartupMessages(library(fda))

  basis <- create.bspline.basis(c(0, 1), 5)
  coef_val <- function(val) matrix(val, 5, 3)

  # Define functional objects
  fd_1 <- fd(coef = coef_val(1), basisobj = basis)
  fd_2 <- fd(coef = coef_val(2), basisobj = basis)

  # Define scalar matrices
  Z1 <- matrix(c(1, 2), nrow = 1)         # 1 × 2
  Z2 <- matrix(c(3, 4), nrow = 1)         # 1 × 2
  Z_many <- matrix(rep(c(1, 2), 3), nrow = 3, byrow = TRUE)  # 3 × 2

  # (1, 1)
  x1 <- predictor_hybrid(Z1, list(fd_1, fd_1), list(matrix(0), matrix(0)), c(5, 5), 1, 2, 2)
  x2 <- predictor_hybrid(Z2, list(fd_2, fd_2), list(matrix(0), matrix(0)), c(5, 5), 1, 2, 2)
  val_scalar <- inprod.predictor_hybrid(x1, x2)

  testthat::expect_type(val_scalar, "double")
  testthat::expect_length(val_scalar, 1)

  # Expected functional inner product: 2 fds × sum(1 * 2) = 2 × (5×3) = 30
  expected_functional <- 2 * sum(1 * 2) * 15  # 5 × 3 = 15 elements
  expected_scalar <- sum(Z1 * Z2)            # 1*3 + 2*4 = 11
  testthat::expect_equal(val_scalar, expected_functional + expected_scalar)

  # (n, 1) broadcasting
  x3 <- predictor_hybrid(Z_many, list(fd_1, fd_1), list(matrix(0), matrix(0)), c(5, 5), 3, 2, 2)
  mat_3_1 <- inprod.predictor_hybrid(x3, x2)
  testthat::expect_equal(dim(mat_3_1), c(3, 1))
  testthat::expect_equal(as.numeric(mat_3_1), rep(val_scalar, 3))

  # (1, m) broadcasting
  x4 <- predictor_hybrid(Z_many, list(fd_2, fd_2), list(matrix(0), matrix(0)), c(5, 5), 3, 2, 2)
  mat_1_3 <- inprod.predictor_hybrid(x1, x4)
  testthat::expect_equal(dim(mat_1_3), c(1, 3))
  testthat::expect_equal(as.numeric(mat_1_3), rep(val_scalar, 3))

  # (n, m)
  mat_3_3 <- inprod.predictor_hybrid(x3, x4)
  testthat::expect_equal(dim(mat_3_3), c(3, 3))
  testthat::expect_true(all(as.numeric(mat_3_3) == val_scalar))

  # shorthand inprod(x) ≡ inprod(x, x)
  self_inner <- inprod.predictor_hybrid(x1)
  testthat::expect_equal(
    self_inner,
    inprod.predictor_hybrid(x1, x1)
  )
})

We check the following cases:
| Case | Test                 | Expectation                             |
| ---- | -------------------- | --------------------------------------- |
| 1    | Orthogonal (Fourier) | Gram matrix ≈ diagonal                  |
| 2    | Constant             | Single value equals interval length (1) |
| 3    | Invalid              | Raises informative error                |
| 4    | Narrow support       | Mostly low off-diagonal entries         |

```{r}
testthat::test_that("Gram matrix edge cases", {
  library(fda)

>>>>

We check the following cases:

| Case | Test                 | Expectation                             |
|------|----------------------|------------------------------------------|
| 1    | Orthogonal (Fourier) | Gram matrix ≈ diagonal                  |
| 2    | Constant             | Single value equals interval length (1) |
| 3    | Invalid              | Raises informative error                |
| 4    | Narrow support       | Mostly low off-diagonal entries         |

```{r}
testthat::test_that("Gram matrix edge cases", {
  suppressPackageStartupMessages(library(fda))

## helpers for functional data objects

### rep_fd

```{r}
#' Replicate a list of single-sample fd objects multiple times
#'
#' This function broadcasts a list of single-sample functional predictors
#' by replicating their coefficient columns.
#'
#' @param fd_list A list of `fd` objects, each representing a single sample.
#' @param n The number of replications desired.
#'
#' @return A list of `fd` objects, each with `n` replications.
#' @export
rep_fd <- function(fd_list, n) {
  if (!is.list(fd_list) || any(!vapply(fd_list, inherits, logical(1), "fd"))) {
    stop("Input must be a list of 'fd' objects.")
  }

  lapply(fd_list, function(fd_obj) {
    coef_mat <- fd_obj$coefs
    if (is.null(dim(coef_mat)) || ncol(coef_mat) != 1) {
      stop("Each fd object in the list must have one column of coefficients.")
    }

    new_coefs <- matrix(rep(coef_mat, n), nrow = nrow(coef_mat), ncol = n)
    fd(coef = new_coefs, basisobj = fd_obj$basis)
  })
}
```

| Case | Description                            | Input Type               | Expected Output                        | Notes                                            |
|------|----------------------------------------|--------------------------|----------------------------------------|--------------------------------------------------|
| 1    | Correct replication                    | List of single-sample fd | List of `fd` objects with `n` columns | All columns identical to original coefficients   |
| 2    | Each fd has one sample                 | List of fd               | No error                               | Verifies structure and content                  |
| 3    | One fd has multiple samples            | fd with >1 columns       | Error                                   | Should raise informative error                  |
| 4    | Non-fd object in list                  | List of non-`fd` items   | Error                                   | Validates class-checking robustness             |

```{r}
testthat::test_that("rep_fd correctly replicates a list of fd objects", {
  suppressPackageStartupMessages(library(fda))

  # Create a simple basis
  basis <- create.bspline.basis(c(0, 1), 5)

  # Create single-sample fd object
  coef_vec <- 1:5
  fd1 <- fd(coef = matrix(coef_vec, ncol = 1), basisobj = basis)

  # Replicate the list
  rep_list <- rep_fd(list(fd1, fd1), 3)

  # Check: list length unchanged
  testthat::expect_equal(length(rep_list), 2)

  # Check each replicated fd object
  for (fd_obj in rep_list) {
    testthat::expect_s3_class(fd_obj, "fd")
    testthat::expect_equal(dim(fd_obj$coefs), c(5, 3))  # 5 basis × 3 samples
    for (i in 1:3) {
      testthat::expect_equal(fd_obj$coefs[, i], coef_vec)
    }
  }

  # Error if not single-sample
  fd_multi <- fd(coef = matrix(1:10, nrow = 5, ncol = 2), basisobj = basis)
  testthat::expect_error(rep_fd(list(fd_multi)), "Each fd object in the list must have one column")

  # Error on non-fd input
  testthat::expect_error(rep_fd(list(1, "not_fd")), "Input must be a list of 'fd' objects")
})

@@

- We define an S3 object `predictor_hybrid` that represent this structure, and implement Hilbert space operations for this class, and our algorithm is built on these foundational operations.

The class definition is a simple list that contains a matrix and a collection of \code{fd} objects from the \code{fda} package, along with the necessary metadata.
```{r classdef:predictor_hybrid}

#' Create a predictor_hybrid object (S3 version, automatic basis size and Gram matrix)
#'
#' Constructs a hybrid predictor object that stores both scalar and functional predictors,
#' and automatically computes Gram matrices (inner products of basis functions).
#'
#' @param Z A numeric matrix of dimension \code{n_sample × n_scalar} representing scalar predictors.
#' @param functional_list A list of functional predictors (e.g., \code{fd} objects from the \code{fda} package).
#'
#' @return An object of class \code{predictor_hybrid}, containing scalar and functional data with Gram matrices.
#' @export
predictor_hybrid <- function(Z, functional_list) {
  stopifnot(is.matrix(Z), is.numeric(Z))
  stopifnot(is.list(functional_list))

@@

  structure(
    list(
      Z = Z,
      functional_list = functional_list,
      gram_list = gram_list,
      n_basis_list = n_basis_list,
      n_sample = n_sample,
      n_functional = n_functional,
      n_scalar = n_scalar
    ),
    class = "predictor_hybrid"
  )
}
```

Let's write some tests to make sure the function behaves as desired:
```{r}
testthat::test_that("predictor_hybrid constructor works as expected (S3, updated)", {
  suppressPackageStartupMessages(library(fda))

@@

  # Check each functional component's coefficients
  expected_coef <- matrix(8, nrow = 5, ncol = 3)
  for (fd_obj in scaled$functional_list) {
    testthat::expect_equal(
      unname(coef(fd_obj)),
      unname(expected_coef)
    )
  }
})
```
### inprod.predictor_hybrid
- The inner product of $f_1=(f_1^{(1)}, \ldots,f_1^{(K)})$ and $f_2=(f_2^{(1)}, \ldots,f_2^{(K)})$ in $\mathcal{F}$ is defined as 
$$\langle f_1, f_2\rangle_\mathcal{F} =  \sum_{k=1}^K \langle f_1^{(k)}, f_2^{(k)}\rangle_{L^2} = \sum_{k=1}^K \int_{\mathcal{T}_k} f_1^{(k)}(t_k) f_2^{(k)}(t_k) dt_k,$$ with norm 
$$\Vert f_1 \Vert_\mathcal{F} = \langle f_1,f_1 \rangle_\mathcal{F}^{1/2} = \{ \sum_{k=1}^K \int_{\mathcal{T}_k} f_1^{(k)}(t_k)^2 dt_k\}^{1/2}.$$
- We define the inner product between any two hybrid objects, $\mathbf{h}_1 = (f_1, \mathbf{v}_1)$ and $\mathbf{h}_2 = (f_2, \mathbf{v}_2)$, as
$$
    \langle \mathbf{h}_1, \mathbf{h}_2\rangle_{\mathcal{H}} = \langle f_1, f_2\rangle_\mathcal{F} +  \langle  \mathbf{v_1}, \mathbf{v_2} \rangle = \sum \limits_{k=1}^K \int_{\mathcal{T}_k} f_1^{(k)}(t_k) f_2^{(k)}(t_k) dt_k + \sum \limits_{r=1}^p v_{1r}v_{2r}, 
$$
with norm $\Vert \cdot \Vert_\mathcal{H} = \langle \cdot,\cdot \rangle_\mathcal{H}^{1/2}$.

- The method `inprod.predictor_hybrid` computes the inner product between two `predictor_hybrid` objects. It supports broadcasting when one of the inputs has a single observation.

@@

  - `xi_1`: A `predictor_hybrid` object.
  - `xi_2`: Another `predictor_hybrid` object. If omitted, defaults to xi_1 (computes self-inner product).

- Value:  A numeric vector of inner products (or a scalar if both inputs are single observations).

- Details:  
  - Functional components are summed using inprod() from the fda package.
  - Scalar components are handled via matrix multiplication.
  - Broadcasting is supported: if either xi_1 or xi_2 has only one sample, its values are broadcast across all rows of the other.
  - Ensures compatibility in the number of functional and scalar predictors before computing the result.
```{r}
#' Inner product between two predictor_hybrid objects (with broadcasting)
#'
#' Computes the inner product between two `predictor_hybrid` objects,
#' including both functional and scalar components. Supports broadcasting
#' when one of the inputs has a single observation.
#'
#' @param xi_1 A `predictor_hybrid` object.
#' @param xi_2 Another `predictor_hybrid` object. If missing, defaults to `xi_1`.
#'
#' @return A numeric vector of inner products, or a scalar if both inputs contain a single observation.
#' @export
inprod.predictor_hybrid <- function(xi_1, xi_2 = NULL) {
  # Handle self-inner product
  if (is.null(xi_2)) xi_2 <- xi_1

  # Type checks
  if (!inherits(xi_1, "predictor_hybrid") || !inherits(xi_2, "predictor_hybrid")) {
    stop("Both inputs must be of class 'predictor_hybrid'.")
  }

  # Structural checks
  if (xi_1$n_functional != xi_2$n_functional) {
    stop("Mismatch in number of functional predictors.")
  }
  if (xi_1$n_scalar != xi_2$n_scalar) {
    stop("Mismatch in number of scalar predictors.")
  }

  # Swap so that broadcasting always applies to xi_2
  if (xi_1$n_sample == 1 && xi_2$n_sample > 1) {
    tmp <- xi_1
    xi_1 <- xi_2
    xi_2 <- tmp
  }

@@

  if (!(n1 == n2 || n2 == 1)) {
    stop("Sample sizes are incompatible for broadcasting.")
  }

  # Prepare components
  f1 <- xi_1$functional_list
  f2 <- xi_2$functional_list
  Z1 <- xi_1$Z
  Z2 <- xi_2$Z

  # Replicate fd and Z if needed
  if (n2 == 1) {
    f2 <- rep_fd(f2, n1)
    Z2 <- matrix(rep(c(Z2), n1), nrow = n1, byrow = TRUE)
  }

  # Compute functional inner products
  inprod_functional <- vapply(seq_len(n1), function(i) {
    sum(vapply(seq_along(f1), function(j) {
      fda::inprod(f1[[j]][i], f2[[j]][i])
    }, numeric(1)))
  }, numeric(1))

  # Compute scalar inner products
  inprod_scalar <- rowSums(Z1 * Z2)

  # Combine results
  result <- inprod_functional + inprod_scalar
  if (n1 == 1 && n2 == 1) as.numeric(result) else result
}
```

We run unit tests for:

| Case | Description                        | Input Dimensions       | Output Length | Notes                                           |
|------|------------------------------------|------------------------|----------------|-------------------------------------------------|
| 1    | Self inner product (1 sample)      | x2: (1×2)              | 1              | Equivalent to squared norm                     |
| 2    | Three-sample vs. Three-sample      | x1: (3×2), x3: (3×2)   | 3              | Dot + functional products, sample-wise         |
| 3    | One-sample vs. One-sample          | x2: (1×2), x4: (1×2)   | 1              | Standard dot + functional product              |
| 4    | Broadcasting (single vs. multi)    | x2: (1×2), x3: (3×2)   | 3              | Broadcast x2 across x3                          |
| 5    | Broadcasting (multi vs. single)    | x3: (3×2), x2: (1×2)   | 3              | Broadcast x2 across x3                          |

```{r}
testthat::test_that("inprod.predictor_hybrid returns correct vector output for broadcasting and other cases", {
  suppressPackageStartupMessages(library(fda))

  basis <- create.bspline.basis(c(0, 1), 5)
  make_fd <- function(val, n_sample) {
    fd(coef = matrix(val, 5, n_sample), basisobj = basis)
  }
# data generation
## three observations
  fd1 <- make_fd(1, 3)
  Z1 <- matrix(c(1, 2), nrow = 3, ncol = 2, byrow = TRUE)
  x1 <- predictor_hybrid(Z1, list(fd1, fd1))
## one observation
  fd2 <- make_fd(2, 1)
  Z2 <- matrix(c(3, 4), nrow = 1)
  x2 <- predictor_hybrid(Z2, list(fd2, fd2))
## three observations
  Z3 <- matrix(c(1, 2, 3, 4, 5, 6), nrow = 3, byrow = TRUE)
  fd3 <- fd(coef = t(matrix(c(1:15), nrow = 3, byrow = TRUE)), basisobj = basis)
  x3 <- predictor_hybrid(Z3, list(fd3, fd3))  
## one observations
  Z4 <- matrix(c(5, 6), nrow = 1)
  fd4 <- make_fd(3, 1)
  x4 <- predictor_hybrid(Z4, list(fd4, fd4))

# Case 1: Self inner product (1 sample)
  out1 <- inprod.predictor_hybrid(x2)
  expected_scalar_1 <- sum(Z2^2)
  expected_functional_1 <- sum(sapply(x2$functional_list, function(fd) fda::inprod(fd, fd)))
  testthat::expect_equal(out1, expected_scalar_1 + expected_functional_1)

# Case 2: Three-sample vs. Three-sample
  out2 <- inprod.predictor_hybrid(x1, x3)
  expected_scalar_2 <- rowSums(Z1 * Z3)
  expected_functional_2 <- vapply(1:3, function(i) {
    sum(sapply(1:2, function(j) {
      fda::inprod(x1$functional_list[[j]][i], x3$functional_list[[j]][i])
    }))
  }, numeric(1))
  testthat::expect_equal(out2, expected_scalar_2 + expected_functional_2)

  # Case 3: One-sample vs. one-sample

  out3 <- inprod.predictor_hybrid(x2, x4)
  expected_scalar_3 <- sum(Z2 * Z4)
  expected_functional_3 <-2*fda::inprod(fd2, fd4)
  testthat::expect_equal(out3, as.numeric(expected_scalar_3 + expected_functional_3))

  # Case 4: Broadcasting (single vs. multi)
  out4 <- inprod.predictor_hybrid(x2, x3)
  expected_scalar_4 <- c(0,0,0)
  for (i in 1:3){
    expected_scalar_4[i] <-  sum(Z2 * Z3[i,])
    }

  expected_functional_4 <-  c(0,0,0)
  for (i in 1:3){
    expected_functional_4[i] <- 2*fda::inprod(
      fd2,
      fd(
        coef = matrix(
        c(
        (5*i+1-5):(5*i)
        ), 5, 1), #end of matrix
         basisobj = basis) #end of fd
      )# end of inprod
  }

  testthat::expect_equal(as.vector(out4), as.vector(expected_scalar_4 + expected_functional_4))

  # Case 4: Broadcasting (multi  vs. single)
  testthat::expect_equal(as.vector(inprod.predictor_hybrid( x3, x2)), as.vector(expected_scalar_4 + expected_functional_4))  
})
```
